---
title: "p8105_hw2_ws2721"
author: "WooJin Shin"
output: github_document
date: "2024-09-30"
---

Problem 1
------------------------------------------------------------
NYC train

```{r}
library(dplyr)
library(tidyr)
```

```{r}
nyc = read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_data.csv", 
               na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  select(line, station_name, station_latitude, station_longitude, 
         contains("route"), entry, vending, entrance_type, ada) |> 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

```{r eval = FALSE}
summary(nyc)
```

```{r eval = FALSE}
nrow(nyc)
```

```{r eval = FALSE}
ncol(nyc)
```

First, I imported the data and converted the variable names to snake cases.
Then, I have selected the variables that I only need and changed the entry variable
to a logical variable.
After the data process, there are 19 columns and 1868 rows.
The data contains 19 variables consists with:
11 character variables (routes 1-7,line, station name, etc.),6 numerical
variables (station latitude, longitude, etc.) and 2 logical variables (entry, ada).

```{r}
nyc2 <- nyc |> 
  mutate(route8 = as.character(route8), route9 = as.character(route9), 
         route10 = as.character(route10), route11 = as.character(route11)) |>
  pivot_longer(cols = route1:route11, names_to = "route", names_prefix = "route",
               values_to = "sub_line") |> 
  drop_na("sub_line")
```

After looking at the questions, to see the distinct stations clearly, I merged the routes to be attached in one station that will make easier for us to isolate the routes.
Also removed the NA to have only lanes that are attached to the station. This will make the dataset more tidy.

```{r eval = FALSE}
station = distinct(nyc2, station_name)

nrow(station)
```

There are 356 distinct stations.

```{r eval = FALSE}
compliant = distinct(nyc2, station_name, ada) |> 
  filter(ada == TRUE)

nrow(compliant)
```

Among 356 distinct stations, 73 are ADA compliant.

```{r eval = FALSE}
allow_entry = filter(nyc2, vending == "NO")|>
  summarise(mean(entry == TRUE))

allow_entry
```

The proportion of station entrances / exits without vending allow entrance is 0.311.

```{r eval = FALSE}
distinct2 = distinct(nyc2, station_name, ada, sub_line) |> 
  filter(sub_line == "A")

nrow(distinct2)
```

57 distinct stations serve A train.

```{r eval = FALSE}
distinct3 = distinct(nyc2, station_name, ada, sub_line) |> 
  filter(sub_line == "A", ada == TRUE)

nrow(distinct3)
```

Out of 57 distinct stations in A train, 16 are ADA compliant.


Problem 2
------------------------------------------------------------
Mr. Trash Wheel

```{r}
library(readxl)

mr_trash = read_excel("./data/202309 Trash Wheel Collection Data.xlsx", 
                   range = "A2:N586", sheet = "Mr. Trash Wheel") |> 
  janitor::clean_names() |> 
  mutate(sports_balls = as.integer(round(sports_balls)),
         year = as.double(year),
         indicator = "Mr. Trash")

prof_trash <- read_excel("./data/202309 Trash Wheel Collection Data.xlsx",
                         range = "A2:M108", sheet = "Professor Trash Wheel") |>
  janitor::clean_names() |> 
  mutate(indicator = "Prof. Trash")

gwynnda_trash <- read_excel("data/202309 Trash Wheel Collection Data.xlsx", 
                      range = "A2:L157", sheet = "Gwynnda Trash Wheel") |>
  janitor::clean_names() |> 
  mutate(indicator = "Gwynnda")

trash_df = bind_rows(mr_trash, prof_trash, gwynnda_trash)
```

I have added the indicator variable for all 3 data sets that I am aggregating and changed the type of variable for Mr. Trash wheel because it gives an error saying that you cannot bind the chr and dbl variables together. So I had to change it to dbl before I bind.

The merged data frame consists of the time, the type and amount of trash gathered in the designated time frame among 3 trash collecting robots. The date is divided to month and year as well as the type of trash (plastic bottles, cigarette_butts .etc) and the amount (weight_tons, volume_cubic_yards .etc). The total amounts of home powered are also included in the data frame.The indicator variable separates the robot in the data frame. 

```{r}
total_prof = trash_df |> 
  filter(indicator == "Prof. Trash") |> 
  summarize(total_weight = sum(weight_tons, na.rm = TRUE))
```

Total collected from Professor Trash Wheel is 216.26 tons.

```{r}
total_cig_gwynnda = trash_df |> 
  filter(indicator == "Gwynnda", 
         month == "June",
         year == "2022") |> 
  summarize(total_cig = sum(cigarette_butts, na.rm = TRUE))
```

Total cigarette butts collected by Gwynnda in June of 2022 is 18120.


Problem 3
------------------------------------------------------------
Great British Bake Off

